---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
### Setting Global options
```{r setoptions, echo=FALSE}
library(knitr)
library(lubridate)
library(lattice)
opts_chunk$set(cache = FALSE)
```


## Loading and preprocessing the data
```{r}
setwd("/Users/shotashimizu//git/Coursera-Data-Science/RepData_PeerAssessment1/")
data <- read.csv("./data/activity.csv")
data$date <- as.Date(data$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r}
stepsSum <- tapply(data$steps, data$date, sum)
hist(stepsSum)
```

The mean is
```{r}
mean(stepsSum, na.rm=T)
```

The median is 
```{r}
median(stepsSum, na.rm=T)
```

## What is the average daily activity pattern?
```{r}
intervalMean <- tapply(data$steps, data$interval, mean, na.rm=T)
plot(intervalMean, type="l")
```

The maximum is 
```{r}
max(intervalMean)
which.max(intervalMean)
```

Imputing missing values
```{r}
sum(is.na(data$steps))
data2 <- data
for(i in seq_along(data2$steps)){
        if(is.na(data2$steps[[i]])){
                a <- as.character(data2$interval[[i]])
                data2$steps[[i]] <- intervalMean[a]
        }
}
```

A histogram for the dataframe with missing values inputted
```{r}
sum(is.na(data2$steps))
head(data2)
stepsSum2 <- tapply(data2$steps, data2$date, sum)
hist(stepsSum2)
```


The mean is
```{r}
mean(stepsSum2, na.rm=T)
```

The median is 
```{r}
median(stepsSum2, na.rm=T)
```

The difference in the mean
```{r}
mean(stepsSum, na.rm=T) - mean(stepsSum2, na.rm=T)
```

The difference in the median
```{r}
median(stepsSum, na.rm=T) - median(stepsSum2, na.rm=T)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
data2$weekend <- (weekdays(data$date,T) == "Sun" | weekdays(data$date,T) == "Sat")
dataWeekend <- with(data2, tapply(steps, list(interval, weekend), mean))
par(mfrow=c(2,1))
plot(dataWeekend[,1], type = "l")
plot(dataWeekend[,2], type = "l")
```
